### 2016-03-16 section/0_intro.tex
split study section. the benchmark study does not guide study, but evaluate study

Figure~\ref{fig:GCC27733} shows a performance bug in GCC. 
Function \texttt{mult\_alg} is a recursive function 
that computes the best algorithm for 
multiplying \texttt{t}, a time-consuming computation. 
At run time, \texttt{mult\_alg} is often invoked for many times, and often
with the same parameter partly due to its recursive nature.
To avoid redundant computation across different instances of
\texttt{mult\_alg} with the same parameter, developers used a
hash-table
\texttt{alg\_hash} to remember which parameter \texttt{t} has been processed
in the past and what is the result. 
Unfortunately, a mistake in the type declaration of hash-table
entry \texttt{hash\_entry} makes the memoization useless for large \texttt{t}.
In many cases, a slow path is taken, when the fast path should have been taken.
This mistake does not affect software correctness, but it causes a large amount
of redundant computation\footnote{This problem will be considered as 
cross-loop redundancy later in this paper, as we consider
recursive functions as a special case of loops.} 
and hurts GCC performance severely,
causing hundreds of times slow down for GCC test cases.

@@

First, figuring out a root-cause taxonomy for inefficient loops.
Such a taxonomy is the prerequisite to developing a general and accurate
diagnosis tool. Guided by a thorough study of 45 real-world inefficient
loop problems, we come up with a hierarchical root-cause taxonomy for
inefficient loops that are
%Specifically, we consider \textit{resultless} and
%\textit{redundancy} as two main categories of root causes for inefficient
%loops, representing loops that spend effort without producing
%any results or producing redundant results. They are further divided into
%sub-categories based on the granularity of the resultless or redundant work. 
general enough to cover
common inefficient loops, and also specific enough to help developers understand
and fix performance problems. More details are
in Section \ref{sec:study}.

@@

We evaluated \Tool on 18 real-world performance problems. 
Evaluation results show that \Tool can accurately identify detailed
root cause for all benchmarks and provide correct fix-strategy suggestion for
16 benchmarks. All of these are achieved with low run-time overhead.

>>>>

Figure~\ref{fig:GCC27733} shows a performance bug in GCC. 
Function \texttt{mult\_alg} is a recursive function 
that computes the best algorithm for 
multiplying \texttt{t}, a time-consuming computation. 
At run time, \texttt{mult\_alg} is often invoked for many times, and often
with the same parameter partly due to its recursive nature.
To avoid redundant computation across different instances of
\texttt{mult\_alg} with the same parameter, developers used a
hash-table
\texttt{alg\_hash} to remember which parameter \texttt{t} has been processed
in the past and what is the result. 
Unfortunately, a mistake in the type declaration of hash-table
entry \texttt{hash\_entry} makes the memoization useless for large \texttt{t}.
In many cases, a slow path is taken, when the fast path should have been taken.
This mistake does not affect software correctness, but it causes a large amount
of redundant computation\footnote{This will be considered as a
loop-redundancy problem later in this paper, as we consider
recursive functions as a special case of loops.} 
and hurts GCC performance severely,
causing hundreds of times slow down for GCC test cases.

@@

First, figuring out a root-cause taxonomy for inefficient loops.
%Specifically, we consider \textit{resultless} and
%\textit{redundancy} as two main categories of root causes for inefficient
%loops, representing loops that spend effort without producing
%any results or producing redundant results. They are further divided into
%sub-categories based on the granularity of the resultless or redundant work. 
Our taxonomy categorizes all inefficient loops into two main categories:
\emph{resultless}, when a large amount of computation produces
no side effect, and \emph{redundancy}, when a large amount of 
computation produces already-available results.
Each main category is then further divided to sub-categories.
We have strived to make our taxonomy both general enough to cover
common inefficient loops, and also specific enough to guide the failure
diagnosis of \Tool, and eventually
help developers understand
and fix performance problems. More details are
in Section \ref{sec:study}.

@@

We evaluated \Tool on \allbugs real-world performance problems,
coming from two representative benchmark suites 
\cite{SongOOPSLA2014,Alabama}. 
Evaluation results show that \Tool can accurately identify detailed
root cause for all benchmarks and provide correct fix-strategy suggestion for
most benchmarks. All of these are achieved with low run-time overhead.

### 2016-03-16 section/1_under.tex
check in proofread results
slight change to figure 3 explanation
all edits for pldi'16 submission
split study section. the benchmark study does not guide study, but evaluate study

This section first presents the root-cause taxonomy designed by us, 
and then discusses the coverage, actionability, and generality
of our taxonomy using a suite of real-world
inefficient loops collected by previous work \cite{SongOOPSLA2014,PerfBug}.

\subsection{Taxonomy Design}
\label{sec:study_tax}
We divide all inefficient loops into two root-cause categories:
\textit{resultless} and \textit{redundancy}.

\subsubsection{Resultless loops}
A resultless loop spends a lot of time in computation without
producing results useful after the loop.
It is further categorized based on which loop iterations are
(not) producing useful results.

@@

{\textbf{[0$|$1]*}}:
Each iteration in this type of loops may or may not produce results.
For some workload, the majority of iterations do not produce
results and cause performance problems perceived by users.
These problems are often solved
by adding a simple check to conditionally skip the expensive loop.
Figure~\ref{fig:GCC46401} shows such an example.
%how users perceive this problem
%explain how every iteration may or may not produce any results
%explain the fix
Users complained that compilation became extremely slow when
the \texttt{-Wsequence-point} checking is enabled.
The slowness was caused
by the \texttt{for} loop in the figure. As the algorithm
behind this loop has quadratic complexity in the number of operands in an expression, 
programs with long expressions suffer severe slow-downs.
After further diagnosis, developers observed that this loop rarely
had any side-effects, as the \texttt{if} condition was rarely satisfied.
At the end, the patch greatly improved performance through a
simple checking that allows the loop to be
skipped for most expressions.

@@

\subsubsection{Redundant loops}
Redundant loops spend a lot of time in repeating computation that is already
conducted. They are further categorized below based on their redundancy unit.

@@

\input section/tab_root

\input section/tab_appbug

\subsection{Taxonomy Assessment}
\label{sec:tax_study}
We assess the coverage, actionability, and generality of our taxonomy using
a set of
real-world inefficient loop problems collected by previous work
\cite{SongOOPSLA2014,PerfBug}.

\subsubsection{Methodology}

Previous work \cite{PerfBug,SongOOPSLA2014} studied the on-line bug
databases of five representative open-source software projects, as 
shown in Table \ref{tab:app_bug}. Through a mix of random sampling and 
manual inspection, they 
found 65 performance problems that are perceived and reported by users. 
Among these 65 problems, 45 problems are related to inefficient loops and 
hence are the target of the study 
here\footnote{The definition of ``loop-related'' in this paper is a little
bit broader than earlier paper~\cite{SongOOPSLA2014}, which only considers
43 problems as loop-related. }.
More details can be found in previous papers that collected
these bugs.

\subsubsection{Assessment}
\label{sec:study_ob}

\comment{
\begin{table*}[tb!]
%\begin{adjustwidth}{-.5in}{-.5in}
\small
\centering
{
\begin{tabular}{|lcccccc|}
\hline
&Apache&Chrome&GCC&Mozilla&MySQL&Total\\
\hline
Total \# of loop bugs & 11 & 4 & 8 & 12 & 10 & 45 \\
\hline
\multicolumn{1}{|l}{Cross-{\bf iteration} Redundancy}
&7&1&2&1&1&12\\
\multicolumn{1}{|l}{ Cross-{\bf loop} Redundancy}
&3&0&2&2&2&9\\
\multicolumn{1}{|l}{ {\bf 0*} Resultness}
&0&0&0&0&0&0\\
\multicolumn{1}{|l}{ {\bf 0*1?} Resultness}
&0&0&0&2&3&5\\
\multicolumn{1}{|l}{{\bf [0$|$1]*} Resultness}
&0&1&1&1&1&4\\
\multicolumn{1}{|l}{{\bf 1*} Resultness}
&1&2&3&6&3&15\\
%&0&2&0&5&0&7&B(4)$|$S(3)\\
%&1&0&3&1&3&8&\\
  %MySQL15811 is moved from 1* to here; consider it as fixed by M
%\hline
%\multicolumn{8}{|c|}{ \# of {\textit {Other}} bugs}\\
%\multicolumn{1}{|l}{Not in above categories}
\hline
\end{tabular}
}
%\end{adjustwidth}
\caption{Number of bugs in each root-cause category. 
}
\label{tab:root}
\end{table*}
}

\comment{
\def\cca#1{\cellcolor{black!#10}\ifnum #1>5\color{white}\fi{#1}}
%For ranges 0-10, multiply by 10 by adding 0 after #1
\begin{table}[tb!]
%\begin{adjustwidth}{-.5in}{-.5in}
\small
\centering
{
\begin{tabular}{lcccccc}
%\hline
                                 &{\bf 0*}    & {\bf 0*1?} & {\bf [0$|$1]*}   & {\bf 1*} & {\bf C-I}  & {\bf C-L} \\
%\hline
 B                               & \cca{0}    & \cca{0}    & \cca{0}          & \cca{4}  & \cca{4}     & \cca{4}   \\
%\hline
 M                               & \cca{0}    & \cca{0}    & \cca{0}          & \cca{0}  & \cca{7}     & \cca{5}   \\
 S                               & \cca{0}    & \cca{1}    & \cca{4}          & \cca{4}  & \cca{0}     & \cca{0}   \\
 C                               & \cca{0}    & \cca{4}    & \cca{0}          & \cca{0}  & \cca{0}     & \cca{0}   \\
 O                               & \cca{0}    & \cca{0}    & \cca{0}          & \cca{7}  & \cca{1}     & \cca{0}   \\
%\hline

\end{tabular}
}
%\end{adjustwidth}
\caption{Number of bugs fixed by each strategy:
B(atching),  
M(emoization), 
S(kipping the loop),
C(hange the data structure), and O(thers). 
``C-I'': cross- iteration redundancy. 
``C-L'': cross-loop redundancy. 
}
\label{tab:correlation}
\end{table}
}

\paragraph{Coverage}
As shown in Table \ref{tab:root}, 
our taxonomy does cover all inefficient loops under study. 
Resultless loops are about as common as redundant loops
(24 vs. 21).
Not surprisingly, 0* loops
are rare in mature software. In fact, no bugs in this
benchmark suite belong to this category.
All other root-cause sub-categories are well represented.

\paragraph{Actionability}
As demonstrated in Table \ref{tab:root}, 
the root-cause categories in our taxonomy are well correlated with
fix strategies.
This indicates that our taxonomy is actionable --- once the root cause
is identified, developers roughly know how to fix the problem.
For example, 
almost all 0*1? resultless loops are fixed by data-structure changes;
all [0$|$1]* resultless loops are
fixed by conditionally skipping the loop;
almost all redundant loops are fixed either by 
memoization or batching. 
In fact, as we will discuss later, simple analysis can tell whether 
memoization or batching should be used to fix a redundant loop.
The only problem is that there are no silver bullets for fixing 1* loops.
%all inefficient loops with xx root cause are fixed by xxx.
%xxx

\paragraph{Generality}
The root-cause categories in our taxonomy are designed to be 
generic. Table \ref{tab:root} also shows that these categories 
each appears in multiple
application in our study. The only exception is $0^*$-resultless, which
never appears.

In summary, the study above informally demonstrates that our taxonomy
is suitable to guide our design of \Tool.
%chances to satisfy the coverage and accuracy requirements of performance 
%diagnosis.
%
%TODO even for the other, there are things we could do ...

\subsubsection{Caveats} 
Just as previous empirical study work, 
our empirical study above needs to be interpreted with our methodology in mind. 
The performance bugs examined above cover a variety of applications, workload, 
development environments and programming languages. 
However, there are still uncovered cases, like distributed systems, 
and scientific computation.

Previous work~\cite{PerfBug, SongOOPSLA2014} uses developers tagging and
on-line discussion to judge whether a bug report is about
performance problems and whether 
the performance problem under discussion is noticed and reported by users
or not.
We follow the methodology used in previous work~\cite{SongOOPSLA2014} to judge 
whether the root-cause of a performance problem is related to loops or not.
We do not intentionally ignore any aspect of loop-related performance problems. 
Some loop-related performance problems may never be noticed by end users
or never be fixed by developers, and hence skip our study. However,
there are no conceivable ways to study them.

\comment{
We believe that the bugs in our study provide a representative sample of the 
well-documented and
fixed performance bugs that are user-perceived and loop-related in the studied 
applications. 
Since we did not set up the root-cause taxonomy to fit particular
bugs in this bug benchmark suite, we believe our taxonomy and diagnosis
framework presented below will go beyond these sampled performance bugs. 
}

>>>>

We divide all inefficient loops into two main root-cause categories,
\textit{resultless} and \textit{redundancy}, and many sub-categories under
these two.
The details of these categories will be presented below.

Our discussion below will qualitatively show that 
our taxonomy is general, not application-specific, and
cover common loop inefficiency problems.
We will quantitatively measure the coverage, actionability, and 
generality of our taxonomy
using real-world performance-bug suite in Section \ref{sec:eval_taxonomy}.

\subsection{Resultless loops}
\label{sec:study_resultless}
A resultless loop spends a lot of time in computation without
producing results useful after the loop.
It is further categorized based on which loop iterations are
(not) producing useful results.

@@

{\textbf{[0$|$1]*}}:
Each iteration in this type of loops may or may not produce results.
For some workload, the majority of iterations do not produce
results and cause performance problems perceived by users.
These problems are often solved
by adding a simple check to conditionally skip the expensive loop.
Figure~\ref{fig:GCC46401} shows such an example.
%how users perceive this problem
%explain how every iteration may or may not produce any results
%explain the fix
Users complained that compilation became extremely slow when
the \texttt{-Wsequence-point} checking is enabled.
The slowness was caused
by the \texttt{for} loop in the figure. As the algorithm
behind this loop has quadratic complexity in the number of operands in an expression, 
programs with long expressions suffer severe slow-downs.
After further diagnosis, developers observed that this loop rarely
had any side-effects, as the \texttt{if} condition was rarely satisfied.
At the end, the patch greatly improved performance through a
simple checking that allows the loop to be
skipped in most cases.

@@

\subsection{Redundant loops}
\label{sec:study_redundant}

Redundant loops spend a lot of time in repeating computation that is already
conducted. They are further categorized below based on their redundancy unit.

@@
